---
title: "Reading List"
subtitle: "Short, recent Goldbach papers (2018-2026)"
date: today
date-format: "MMMM YYYY"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(DT)
papers <- read.csv("../out-oa/gb_short_papers_v04.csv",
                   stringsAsFactors = FALSE,
                   fileEncoding = "UTF-8")
```

A curated list of short Goldbach-relevant papers, sorted shortest first. The intended audience is a strong undergraduate or motivated high school student who wants to read recent research without committing to a 60-page monograph.

::: {.content-block}
**How the list was built.** OpenAlex was queried for all papers matching `Goldbach conjecture`, `Goldbach problem`, or `Goldbach's conjecture` published since 2018. Filters applied: at most 10 authors (no physics megapapers), at most 15 pages when page count is known. Sort: shortest first, then most recent, then most cited.
:::

`r nrow(papers)` papers in total. Click any title to open the full paper.

```{r reading-table}
display <- papers %>%
  mutate(
    title_link = ifelse(nzchar(best_url),
                        sprintf('<a href="%s" target="_blank">%s</a>',
                                best_url, title),
                        title),
    pages_disp = ifelse(is.na(pages) | pages == "", "?", as.character(pages))
  ) %>%
  transmute(
    `#`     = rank_placeholder,
    Pages   = pages_disp,
    Year    = year,
    Cites   = cites,
    Title   = title_link,
    Authors = substr(authors, 1, 80)
  )

DT::datatable(
  display,
  rownames = FALSE,
  escape = FALSE,
  options = list(
    pageLength = 50,
    lengthMenu = c(25, 50, 100, 200),
    order = list(list(1, 'asc')),
    dom = 'fltip',
    columnDefs = list(
      list(className = 'text-right', targets = c(0, 1, 2, 3))
    )
  ),
  class = 'compact stripe hover'
)
```
