---
title: "The Top 100"
subtitle: "Researchers active on Goldbach and adjacent problems"
date: today
date-format: "MMMM YYYY"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(DT)
top100 <- read.csv("../out/gb_top100_v02.csv",
                   stringsAsFactors = FALSE,
                   fileEncoding = "UTF-8")
```

The list is sortable on any column. Type in the search box to filter by name, institution, or country.

```{r top100-table}
display <- top100 %>%
  transmute(
    Rank        = merged_rank,
    Name        = name,
    Institution = institution,
    Country     = country,
    `arXiv rank` = ifelse(arx_rank == 156, "-", as.character(arx_rank)),
    `OA rank`   = ifelse(oa_rank == 1114,  "-", as.character(oa_rank)),
    Source      = provenance,
    `First year` = first_year,
    `Last year`  = last_year
  )

DT::datatable(
  display,
  rownames = FALSE,
  options = list(
    pageLength = 25,
    lengthMenu = c(25, 50, 100),
    order = list(list(0, 'asc')),
    dom = 'fltip'
  ),
  class = 'compact stripe hover'
)
```

## Provenance breakdown

* **`both`**: in arXiv 155 AND OpenAlex top pool. The unanimous core.
* **`oa_only`**: discovered by OpenAlex but not arXiv-active. These are the senior figures who publish in journals and don't post preprints.
* **`arxiv_only`**: present only if at least one arXiv-only entry made the sum_rank cutoff (currently 0 of 100; sum penalty for not appearing in OA is large).
* **`manual_add`**: manually inserted via `gb_manual_additions.csv` after auditor review.

```{r provenance}
prov <- top100 %>% count(provenance, name = "n") %>% arrange(desc(n))
knitr::kable(prov, col.names = c("Source", "Count"), align = c("l", "r"))
```

## Limitations

This list reflects publication output and citation network density, not subjective importance. A working theorist whose entire body of Goldbach work is one influential paper may rank lower than a productive researcher with many adjacent papers. For historical perspective and active-research-front, this list is a starting point, not a verdict.
